### 2016-06-07 paper/icfp16-cameraready/related.tex
misc

% \subparagraph{Repairing}
In addition to localizing the error, \citet{Lerner2007-dt} attempt to
suggest a fix by replacing expressions (or removing them entirely) with
alternatives based on the surrounding program context.
%
\citet{Chen2014-gd} use a variational type system to allow for the
possibility of changing an expression's type, and search for an
expression whose type can be changed such that type inference would
succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types used in the program.
%
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}
%

@@

\paragraph{Program Exploration}
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} --- containing holes --- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera's system focuses on backward exploration; in
contrast, our visualization supports forward \emph{and} backward
exploration, though our backward steps are more limited than
Perera's. Specifically, we do not support selecting a value and
inserting the intermediate terms that preceded it while ignoring
unrelated computation steps. %; this would be interesting future work.

>>>>

% \subparagraph{Repairing}
In addition to localizing the error, \citet{Lerner2007-dt} attempt to
suggest a fix by replacing expressions (or removing them entirely) with
alternatives based on the surrounding program context.
%
\citet{Chen2014-gd} use a variational type system to allow for the
possibility of changing an expression's type, and search for an
expression whose type can be changed such that type inference would
succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner~\etal, who search for changes at the value-level,
Chen~\etal\ search at the type-level and are thus complete due the finite
universe of types used in the program.
%
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}
%

@@

\paragraph{Program Exploration}
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} --- containing holes --- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera~\etal\ focus on backward exploration; in contrast, our
visualization supports forward \emph{and} backward exploration, though
our backward steps are more limited.
%
Specifically, we do not support selecting a value and inserting the
intermediate terms that preceded it while ignoring unrelated computation
steps. %; this would be interesting future work.

### 2016-06-07 paper/icfp16-cameraready/semantics.tex
clarify various things relating to language features and inhabitants
misc

\begin{thm}{\textbf{[Witness Generality]}}
\label{thm:soundness}
  For any lambda-term $f$, if
  \hbox{$\steptr{\trace}{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu}{\stuck}{\vsu}{\tsu}$,}
  then for every % inhabitable
  type
  % \footnote{We exclude builtin functions that subvert the type system, \eg \texttt{Obj.magic}, and thus consider the type $\forall a b. \tfun{a}{b}$ to be uninhabitable.}
  $t$ there exists a value $v$ of type $t$ such that
  $\steps{\eapp{f}{v}}{\emptysu}{\emptysu}{\stuck}{\vsu'}{\tsu'}$.
\end{thm}

@@

\paragraph{Inputs and Outputs}
%
The problem of generating inputs is undecidable in general.
%
Our witness generation procedure takes two inputs.
%
First, a search bound $k$ which is used to define the \emph{number} of
traces to explore. (We assume, without loss of generality, that all
traces are finite.)
%
Second, the target expression $e$ that contains the type error,
and may be a curried function of multiple arguments.
%
The witness generation procedure returns as output a list of (general)
witness expressions, each of which is of the form $e\ v_1 \ldots v_n$.
%
The \emph{empty} list is returned when no witness can be found after
exploring $k$ traces.

>>>>

\begin{thm}{\textbf{[Witness Generality]}}
\label{thm:soundness}
  For any lambda-term $f$, if
  \hbox{$\steptr{\trace}{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu}{\stuck}{\vsu}{\tsu}$,}
  then for every
  (inhabited\footnote{All types in our formalism are inhabited, but in a larger language like \ocaml this may not be the case.})
  type
  % \footnote{We exclude builtin functions that subvert the type system, \eg \texttt{Obj.magic}, and thus consider the type $\forall a b. \tfun{a}{b}$ to be uninhabitable.}
  $t$ there exists a value $v$ of type $t$ such that
  $\steps{\eapp{f}{v}}{\emptysu}{\emptysu}{\stuck}{\vsu'}{\tsu'}$.
\end{thm}

@@

\paragraph{Inputs and Outputs}
%
The problem of generating inputs is undecidable in general.
%
Our witness generation procedure takes two inputs:
%
(1) a search bound $k$ which is used to define the \emph{number} of
traces to explore\footnote{We assume, without loss of generality, that all
traces are finite.} and
%
(2) the target expression $e$ that contains the type error
(which may be a curried function of multiple arguments).
%
The witness generation procedure returns as output a list of (general)
witness expressions, each of which is of the form $e\ v_1 \ldots v_n$.
%
The \emph{empty} list is returned when no witness can be found after
exploring $k$ traces.
